---
title: "Lupus App 2.0"
author: "Itzel Olivares"
toc: true # Activate table of contents
toc-depth: 3 # Table of contents depth
toc-expand: true # Show table expanded at the beginning
toc-location: left-body # Table of contents location
number-sections: true # Number headings in the render
number-depth: 3 # Number headings up to level 3
toc-title: "Contents" # Custom TOC title
theme: flatly # Theme
fontsize: 100% # Base font size
# fontcolor: grey # Optional text color
highlight-style: pygments
format:
  html:
    embed-resources: true
    code-fold: true
    html-math-method: katex
  pdf:
    geometry:
      - top=30mm
      - left=30mm
  docx: default
editor:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
# Packages used in the project
required_packages <- c(
  "tidyverse",
  "visdat",
  "janitor",
  "skimr",
  "readr",
  "lubridate",
  "stringr",
  "knitr",
  "kableExtra")

# Feature for automatically installing and downloading packages
load_or_install <- function(packages) {
  for (pkg in packages) {
    if (!require(pkg, character.only = TRUE)) {
      install.packages(pkg, dependencies = TRUE)
      library(pkg, character.only = TRUE)
    }
  }
}

load_or_install(required_packages)

```

# Initial Workflow

1.  Write an introduction for the Mexican Lupus Registry App

2.  Describe study population

3.  Start the curation process as follows:

    1.  Preserve the original REDCap variable names to ensure
        consistency and traceability across all stages of the project.

    2.  Maintain the participant identifier (record_id) unchanged to
        allow accurate data linkage.

    3.  Select variables based on expert consensus derived from the
        Delphi process.

    4.  Create a curated dataset that includes only the selected
        variables.

4.  Evaluate missing data patterns using visualization techniques,
    without excluding subjects or variables.

5.  Prepare the dataset for downstream descriptive analysis, plotting, and statistical modeling.

6.  In a second stage, integrate the Neurolupus dataset—containing MRI findings, VolBrain metrics, and cognitive assessments (MoCA)—using the same participant identifier.

# Introduction

::: {style="text-align: justify;"}
This document presents the set of variables selected for the project
“Building Accessible Resources to Empower Communities: The Case of the Mexican Lupus Registry.” The project aims to develop an interactive application that enables users to explore data from the Mexican Lupus Registry, generate descriptive statistical reports, and build predictive models based on user-defined combinations of variables.

The selection of these variables is designed to balance clinical
relevance, data availability, and analytic utility, ensuring that the application supports both research and patient-centered insights.
:::

# Methods

## Study Population and Data Source

::: {style="text-align: justify;"}
We analyzed data from the LupusRGMX cohort <https://www.frontiersin.org/journals/digital-health/articles/10.3389/fdgth.2024.1344103/full>, an electronic registry of Mexican patients with Systemic Lupus Erythematosus (SLE) <https://doi.org/10.1371/journal.pdig.0000706>. For this analysis, we included patients aged 18 years or older with a confirmed diagnosis of SLE. Data curation involved the removal of duplicate records and the exclusion of variables with 100% missing values.
:::

## Variable selection process

::: {style="text-align: justify;"}
A committee of six experts in the LupusRGMX dataset cohort was formed, and they carried out the variable selection process using a consensus-based approach.

1.  The experts independently reviewed the complete list of variables and selected 20 variables, providing justification for their selection.

2.  A list was created in which the variables were ranked from highest to lowest concordance, based on each expert’s selection.

3.  A meeting was held to discuss each variable on the list, and a vote was taken to determine whether the variable in question would be retained or eliminated

4.  Variables that were clinically relevant and aligned with the intended use of the application were retained, as were those useful for the planned analytic and reporting modules.

Once selected, they were grouped into: 

- Clinical variables 
- Sociodemographic variables 
- PROMs (Patient - Reported Outcomes) 
- Lifestyle and quality of life 
- Neurolupus

The final curated dataset must retain the original names of the RedCap variables, keep the participant identifier intact, and avoid removing subjects or variables during the curation stage.
:::

## Selected Variables and Coding

::: {style="text-align: justify;"}
All variables were preserved using their original REDCap names to ensure traceability and reproducibility. Variables were classified according to their data type as continuous, binary, or categorical (nominal or ordinal), and coding schemes were maintained according to the original data dictionary.
:::

::: {style="text-align: justify;"}
Selected variables were organized into clinically meaningful domains, including demographic characteristics, disease activity and organ involvement, treatment, reproductive health, lifestyle factors, and patient-reported outcomes. Coding schemes were maintained according to the original REDCap data dictionary. *(Add reference to dictionary if possible)*
:::

### Demographic and Socioeconomic Variables

::: {style="text-align: justify;"}
Demographic variables included:

- participant identification (`record_id`)
- sex (coded as binary checkbox variables `sex___1==1` for female and `sex___1==0` for male), and 
- age (`calculated_age`, continuous).

Geographic variables such as place of birth (`place_of_birth`) and
current residence (`home`) were coded according to the 32 federal
entities of Mexico published by INEGI.

Educational attainment (`school`) and occupation (`ocupation`) were
coded as ordinal categorical variables reflecting increasing levels of
education and employment status. 

Specifically, `ocupation` was coded as
1 = Student
2 = Employed
3 = Unemployed
4 = Retired. 

The variable `school` was coded as 
1 = None
2 = Primary education
3 = Secondary education
4 = High school or technical degree
5 = Bachelor’s degree
6 = Postgraduate education.

Socioeconomic status (`socioeconomic_status`) was included as a
composite index following  the rule (NSE 2022) of the Mexican Association of Market and Opinion Intelligence Agencies (AMAI) <https://www.amai.org/NSE/index.php?queVeo=NSE2020>.

:::

### Lifestyle: Dietary pattern and exercise

::: {style="text-align: justify;"}
Dietary behavior was assessed using the Eating Behavior Questionnaire,
incorporating both a total score (`puntuacion_alimentacion`) and a
categorical classification (`clasi_ali`). The total score reflects
overall eating patterns and was used to classify dietary habits into
three categories.

- A score of 3 corresponds to adequate dietary habits, indicating balanced
eating patterns, appropriate distribution of food groups, and regular
meal timing, including breakfast, lunch, dinner, and snacks. This
category reflects individuals who maintain a structured and balanced
diet, which is associated with better overall health and daily
functioning.

- A score of 2 indicates partially inadequate dietary habits, reflecting
mild imbalances in food intake or meal timing. These individuals may
exhibit inconsistencies in dietary patterns, such as irregular meals or
unbalanced consumption across food groups, which may impact nutritional
quality.

- A score of 1 corresponds to inadequate dietary habits, indicating a lack
of balance in food intake, including excessive consumption of sugars,
saturated fats, or salt, as well as irregular or poorly structured meal
patterns. This category reflects dietary behaviors that may negatively
impact health and increase the risk of disease.

Overall, higher scores are indicative of healthier and more balanced
dietary patterns.

Physical activity was assessed using the International Physical Activity
Questionnaire (IPAQ), with total physical activity summarized through
the variable `metstotal`, representing the overall metabolic equivalent
task (MET) score. This variable was treated as a continuous measure of
energy expenditure.

Additionally, `metstotal` was interpreted using standardized thresholds:

- values below 600 MET-minutes/week were classified as low physical activity, indicating insufficient intensity or duration to meet moderat or high activity criteria. 

- Scores equal to or greater than 600 MET-minutes/week were considered indicative of moderate physical activity, defined by patterns such as at least 3 days of vigorous activity for a minimum of 20 minutes per day, 5 days of moderate
activity and/or walking for at least 30 minutes per day, or equivalent
combinations. 

- Scores exceeding 1500 MET-minutes/week were classified as high physical activity, corresponding to sustained engagement in moderate and/or vigorous activities across most days of the week.
:::

### Clinical Variables

::: {style="text-align: justify;"}
Family history of lupus was assessed using:

- `family_sle`, coded as 0 = No and 1 = Yes

Aand further detailed through checkbox variables identifying
affected relatives (`family_member_sle___0` to `family_member_sle___8`),
each coded as a binary indicator (1 = Yes, 0 = No). These variables
correspond to specific family members as follows:

- `family_member_sle___0` = Father
- `family_member_sle___1` = Mother
- `family_member_sle___2` = Siblings
- `family_member_sle___3` = Grandparent
- `family_member_sle___4` = Aunt/Uncle
- `family_member_sle___5` = Cousin
- `family_member_sle___6` = Niece/Nephew
- `family_member_sle___7` = Child
- `family_member_sle___8` = Other

Temporal variables included age at diagnosis (`dx_age`) and time since diagnosis (`dx_time`), both treated as continuous variables ranging from 0 to 99 years. 

Time from symptom onset to diagnosis (`time_symptoms`) was captured as an ordinal categorical variable reflecting increasing duration intervals, coded as follows: 

0 = less than 6 months
1 = between 6 months and 1 year
2 = 1 year
3 = 2 years
4 = 3 years
5 = 4 years
6 = more than 5 years
7 = more than 10 years

Anthropometric measurements included weight (`weight`), height (`height`), body mass index (`bmi`), waist circumference (`waist`), and hip circumference (`hips`), all treated as continuous variables.

Comorbidities were captured using the variable `comorbidities`, which
includes predefined categories coded as follows: 

1 = Diabetes mellitus
2 = Hypertension 
3 = Cancer
4 = Cardiovascular disease
5 = Osteoporosis
6 = Rheumatoid arthritis
7 = Multiple sclerosis
8 = Thyroid disorders
9 = Other
11 = None

This variable allows the identification of coexisting conditions relevant to disease burden and clinical outcomes.

Cardiovascular risk was evaluated using derived indicators such as
estimated cardiovascular risk (`ht_cvd_risk`) and heart age
(`ht_heart_age`), both treated as continuous variables.

Clinical characterization included measures of disease activity and
cumulative damage, such as the total SLEDAI score (`total sledai`) and
the SLICC damage index (`total slicc`), both of which were treated as
continuous variables. The disease activity classification
(`Disease Activity Classification`) was included as a categorical
variable, with a score of \>6 indicating active disease.

Organ involvement was represented by the domains included in the SLICC
damage index. These included 

renal (`Renal`)
cardiovascular (`Cardiovascular`)
pulmonary (`Pulmonary`)
gastrointestinal (`Gastrointestinal`)
musculoskeletal (`Musculoskeletal`)
neuropsychiatric (`Neuropsychiatric`)
peripheral vascular (`Peripheral vascular`)
cutaneous (`Skin`)
ocular (`Ocular`)
cardiomyopathy (`Cardiomyopathy`)
diabetes (`Diabetes`)
malignancy (`Malignancy`). 

These variables were coded as binary indicators (1 =
present, 0 = absent).
:::

### Treatment Variables

::: {style="text-align: justify;"}
Pharmacological treatment variables included corticosteroid use
(`prednisolona`), coded according to the specific drug type as follows:

0 = Prednisone
1 = Prednisolone
2 = Deflazacort
3 = Meticorten
4 = Methylprednisolone
5 = Calcort
6 = Betamethasone
7 = No corticosteroid use
8 = Other. 

Dosage was captured using
`pred_dosis`, measured in milligrams (mg), and further categorized using
`pred_dosis_categ` into low and high dose groups.

Immunosuppressive and adjunct therapies were represented using checkbox
variables (`treatment_0` to `treatment_8`), where each category
corresponds to a specific medication class and is coded as a binary
indicator (1 = Yes, 0 = No). 

These categories include: 
`treatment_0` = Antimalarials (hydroxychloroquine, chloroquine)
`treatment_1` = Corticosteroids (prednisone, prednisolone, deflazacort, meticorten, methylprednisolone, calcort, betamethasone), `treatment_2` =
Mycophenolic acid (mycophenolate)
`treatment_3` = Azathioprine
`treatment_4` = Methotrexate 
`treatment_5` = Rituximab
`treatment_6` = Cyclosporine
`treatment_7` = Cyclophosphamide
`treatment_8` = Other treatments.

Additional treatment information was captured through
`aditional_treatment`, coded as a binary variable (1 = Yes, 2 = No),
allowing further characterization of therapeutic regimens.
:::

### Reproductive Health Variables

::: {style="text-align: justify;"}
Reproductive history variables included current pregnancy status
(`pregnant`), coded as 1 = Yes and 0 = No, as well as pregnancy planning
(`pregnancy_plan`), defined as intention to conceive in the near future
(1 = Yes, 0 = No). 

Previous pregnancy history was captured through
`pregnancies` (1 = Yes, 0 = No)
`pregnancies_number` (total number of pregnancies, ranging from 1 to 5)
`pregnancies_2`, indicating alloimmunization status (1 = Yes, 0 = No).

Detailed obstetric history was assessed using grouped variables
describing pregnancy outcomes (`pregn_end`, `pregn_end_2`,
`pregn_end_3`, `pregn_end_4`), coded as 0 = Term, 1 = Preterm, and 2 =
Miscarriage. 

Pregnancy complications were captured using `pregn_compl`
and related variables, where 0 = Preeclampsia, 1 = Eclampsia, and 2 =
HELLP syndrome. 

Timing of complications (`pregn_compl_time` and related
variables) was recorded according to gestational month, ranging from 1
(first month) to 9 (ninth month). 

Preeclampsia history (`preeclampsie` and related variables) was evaluated for each pregnancy and coded as 0 = No, 1 = Yes, and 2 = Unknown.

Lifestyle factors during each pregnancy included smoking status
(`smoking`), coded as 1 = Yes and 0 = No. 
Breastfeeding history was captured using `breastfeeding` (0 = No, 1 = Yes) and `breasfeeding_type` variables, specifying type per pregnancy as 1 = Exclusive, 2 = Partial, and 3 = None or not applicable.

Menstrual and gynecological history variables included age at menarche
(`period_age`, in years), presence of menstruation in the past 12 months
(`last_period)`, 0 = No, 1 = Yes, and age at last menstruation
(`last_period_age`, in years). Cycle regularity (`cycle_regularity`) was
coded as 0 = No and 1 = Yes. Surgical history included hysterectomy
(`histerectomy`) and oophorectomy (`ooforectomy`), both coded as 0 = No
and 1 = Yes, with `ooforectomy_ovaries` specifying the number of ovaries
removed (1 = one ovary, 2 = both ovaries, 3 = unknown).

Contraceptive use was evaluated through current use (`contraceptives`, 1
= Yes, 0 = No) and type (`contraceptive_type`), categorized as 

1 = Combined (estrogen/progestin)
2 = Progestin-only
3 = Unknown
4 = Other. Past contraceptive use was captured through
`contraceptives_past`, age at use (`contraceptive_past_age`), duration
(`contraceptives_past_time`, coded as 0 = \<1 year and 1 = ≥1 year), and
type (`contraceptive_past_type`) using the same categories.

Menopause-related variables included hormone therapy use
(`menopause_tx`, 0 = No, 1 = Yes) and type of therapy
(`menopause_tx_type`), categorized as 

1 = Combined (estrogen/progestin)
2 = Progestin-only
3 = Unknown
4 = Other
:::

### Patient-Reported Outcomes and Psychological Measures

::: {style="text-align: justify;"}
Patient-reported outcomes included measures of quality of life, sleep
quality, pain, mental health and psychological characteristics like
resilience and personality.

Global quality of life was assessed using `cal_totalcv`, derived from
the WHOQOL instrument, where scores greater than 40 indicate high
quality of life, reflecting individuals with active physical and mental
engagement, positive social relationships, and a favorable perception of
their position within their cultural and value systems.

Sleep quality was evaluated using `cal_glob`, a composite score based on
the Pittsburgh Sleep Quality Index (PSQI), which assesses sleep quality
and related disturbances over the previous month. Scores range from 0 to
21, with higher values indicating poorer sleep quality. Specifically,

scores \<5 suggest no sleep problems
5–7 indicate the potential need for medical attention
8–14 suggest the need for clinical evaluation and possible treatment
≥15 are indicative of severe sleep disturbances.

Mental health was assessed using `calc_psicologico`, a WHOQOL-derived
domain score (maximum of 24 points, where higher values indicate greater
satisfaction), and `salud_mental`, an SF-36 domain score ranging from 0
to 100, where higher scores reflect better perceived mental health
status.

Psychological symptomatology was assessed using the SCL-90-R instrument,
with domain-specific continuous scores including 

somatization (`scl90r_score_som`, range 11–55)
obsessive-compulsive symptoms (`scl90r_score_oc`, range 7–35)
interpersonal sensitivity (`scl90r_score_si`, range 7–35)
depression (`scl90r_score_dep`, range 12–60)
anxiety (`scl90r_score_ans`, range 9–45)
hostility (`scl90r_score_hos`, range 5–25)
phobic anxiety (`scl90r_score_fob`, range 3–15)
paranoid ideation (`scl90r_score_para`, range 5–25)
psychoticism (`sci90r_score_psi`, range 7–35). 

Higher scores across these domains are associated with greater symptom severity, including somatic complaints, emotional distress, maladaptive thought patterns, and functional impairment.

Physical pain was evaluated using `dolor_corporal_trans`, an SF-36
transformed domain score ranging from 0 to 100, where higher values
indicate better health status and lower perceived pain.

Resilience was assessed using the Mexican Resilience Scale, which
includes multiple domain-specific scores: 

`factunofcmsumatotal` (strength and self-confidence)
`factdoscompsumatotal` (social competence)
`factresapoyofamsumatotal` (family support)
`factcuatroapoyosocsumt` (social support)
`factcincoestructsumtotal` (structure)

Higher scores indicate greater resilience across domains. Specifically, 

scores \>38 in strength and self-confidence reflect strong goal clarity, perseverance, and optimism

scores \>16 in social competence indicate effective coping and adaptability

scores \>12 in family support reflect strong familial
bonds and shared values

scores \>10 in social support indicate the presence of reliable social networks; 

scores \>10 in structure reflect the ability to organize and plan effectively even in challenging situations.

Personality traits were evaluated using the Five-Factor Model (Big
Five), including 

extroversion (`big_five_extro`)
agreeableness (`big_five_agra`)
conscientiousness (`big_five_cons`)
neuroticism (`big_five_neurot`)
openness to experience (`big_five_apert`)

Higher scores in extroversion (\>20) are associated with sociability and enthusiasm

higher agreeableness (\>22) reflects altruism and cooperativeness

higher conscientiousness (\>22) indicates organization and self-discipline

higher neuroticism (\>20) is associated with emotional instability and vulnerability to stress

higher openness (\>25) reflects curiosity, creativity, and openness to new experiences.

Lower scores in each domain correspond to the inverse behavioral
tendencies.
:::

::: {style="text-align: justify;"}
All variables retained their original REDCap structure, including
checkbox encoding (suffix `___n`), categorical coding, and calculated
fields. No recoding or renaming was performed, ensuring full alignment
with the original data dictionary and facilitating reproducible analysis
workflows.
:::

# Reading data set

```{r}
#| label: Reading data set
#| include: false

db.01 <-read.csv("~/R/Shinnyapp/base_de_datos_sin_registros_duplicados_ LupusProjectProducti_DATA_2026-01-06_1544.csv sin_col_vacias_con_suma_SLICC_SLEDAI_pred_dosis_categ_dx_time_curado (1).csv")
```

## Missingness check

```{r Selecting variables, fig.width=10, fig.height=6}
# Select variables
db.01_selected <- db.01 %>%
  dplyr::select(
     record_id, 
     sex___1, 
     #sex___2, 
     calculated_age, 
     place_of_birth, 
     dx_age,
     prednisolona, 
     pred_dosis, 
     comorbidities, 
     dx_time, time_symptoms,
     treatment___0, 
     treatment___1, 
     treatment___2, 
     treatment___3,
     treatment___4, 
     treatment___5, 
     treatment___6, 
     treatment___7,
     treatment___8, 
     aditional_treatment, 
     family_sle,
     family_member_sle___0, 
     family_member_sle___1, 
     family_member_sle___2,
     family_member_sle___3, 
     family_member_sle___4, 
     family_member_sle___5,
     family_member_sle___6, 
     family_member_sle___7, 
     family_member_sle___8,
     home, 
     ocupation, 
     school, 
     socioeconomic_status,
     calc_psicologico, 
     cal_totalcv, 
     dolor_corporal_trans,
     salud_mental_trans, 
     cal_glob,
     scl90r_score_som, 
     scl90r_score_oc, 
     scl90r_score_si,
     scl90r_score_dep, 
     scl90r_score_ans, 
     scl90r_score_hos,
     scl90r_score_fob, 
     scl90r_score_para, 
     sci90r_score_psi,
     factunofcmsumatotal, 
     factdoscompsumatotal, 
     factresapoyofamsumatotal, 
     factcuatroapoyosocsumt,
     factcincoestructsumtotal,
     big_five_extro, 
     big_five_agra, 
     big_five_cons, 
     big_five_neurot, 
     big_five_apert, 
     height, 
     weight, 
     bmi, 
     waist, 
     hips,
     metstotal, 
     puntuacion_alimentacion, 
     clasi_ali,
     pregnant, 
     pregnancy_plan, 
     pregnancies, 
     pregnancies_number,
     pregnancies_2, 
     pregn_end, 
     pregn_end_2, 
     pregn_end_3, 
     pregn_end_4,
     preeclampsie, 
     pregn_compl, 
     pregn_compl_time,
     preeclampsie_2, 
     pregn_compl_2, 
     pregn_compl_time_2,
     preeclampsie_3, 
     pregn_compl_3, 
     pregn_compl_time_3,
     preeclampsie_4, 
     pregn_compl_4, 
     pregn_compl_time_4,
     smoking, 
     smoking_2, 
     smoking_3, 
     smoking_4,
     breastfeeding, 
     breasfeeding_type, 
     breasfeeding_type_2,
     breasfeeding_type_3, 
     breasfeeding_type_4,
     period_age, 
     last_period, 
     last_period_age,
     cycle_regularity, 
     histerectomy, 
     ooforectomy,
     ooforectomy_ovaries,
     contraceptives, 
     contraceptive_type,
     contraceptives_past, 
     contraceptive_past_age,
     contraceptives_past_time, 
     contraceptive_past_type,
     menopause_tx, 
     menopause_tx_type,
     ht_heart_age, 
     ht_cvd_risk,
     Ocular, 
     Neuropsychiatric, 
     Renal, 
     Pulmonary,
     Cardiovascular, 
     Cardiomyopathy, 
     Peripheral.vascular,
     Gastrointestinal, 
     Musculoskeletal,  
     Skin,
     Premature.gonadal, 
     Diabetes, 
     Malignancy,
     total.slicc, 
     total.sledai, 
     Disease.activity.classification,
     pred_dosis_categ
     )
```

```{r vismiss_plot_01, fig.width=10, fig.height=6}
#View missing data by groups, from least to most missing
#View data for variables 1–30 
visdat::vis_miss(db.01_selected[, 1:30])
```

```{r vismiss_plot_02, fig.width=10, fig.height=6}
#View data for variables 31–60
visdat::vis_miss(db.01_selected[, 31:60])
```

```{r vismiss_plot_03, fig.width=10, fig.height=6}
#View data for variables 61–90
visdat::vis_miss(db.01_selected[, 61:90])
```

```{r vismiss_plot_04, fig.width=10, fig.height=6}
#View data for variables 91–130
visdat::vis_miss(db.01_selected[, 91:130])
```

Let's see the most missed variables

```{r Most_missed_variables}
db.missed <- db.01_selected %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "na_count") %>%
  mutate(na_percent = na_count / nrow(db.01_selected) * 100) %>%
  arrange(desc(na_count))
```


::: {style="text-align: justify;"}
The visualization of missing data reveals a high proportion of incomplete values in specific domains, particularly within reproductive health variables. Given the exploratory nature of this stage, no variables or observations were removed. However, for downstream analyses, it is recommended to consider reducing dimensionality by excluding reproductive variables with high missingness or prioritizing those with greater clinical relevance and completeness.
:::

### Selecting most complete variables

After checking NAs, we established an threshold of 1500, in other words, variables with less than 1500 NAs was kept.

```{r}
#| label: Selcting variables with less than 1500 NAs

vars_keep <- db.missed %>%
  filter(na_count < 1500) %>%
  pull(variable)

db.03 <- db.01_selected %>%
  select(all_of(vars_keep)) %>% 
  na.omit()

visdat::vis_miss(db.03)
```

### Check data structure

```{r Checkin data structure}
glimpse(db.03)
```

### Exporting data set

```{r}
#| label: Exporting selected variables
write_csv(db.03, "Lupus_App_dataset.csv")

# Verifying dataset 
db.04 <- read.csv("Lupus_App_dataset.csv")
glimpse(db.04)
```


# Second stage: neurolupus integration

The second stage integrates the neurolupus information with the curated registry data. This includes the radiology interpretation dataset, the Volbrain outputs, and the MoCA assessment. The merge must be performed using the same RedCap participant ID.

```{r neuro_integration}
# Here is where the neurolupus file will be added

```
