---
title: "Lupus App 2.0,
        Building accessible resources to empower communities: 
        the case of the Lupus Mexican Registry"
author: "Dris Sánchez, 
         Anita Ledesma,
         Grecia Sevilla,
         Fernanda Bravo,  
         Itzel Olivares,
         Domingo Martínez"
toc: true # Activate table of contents
toc-depth: 3 # Table of contents depth
toc-expand: true # Show table expanded at the beginning
toc-location: left-body # Table of contents location
number-sections: true # Number headings in the render
number-depth: 3 # Number headings up to level 3
toc-title: "Contents" # Custom TOC title
theme: flatly # Theme
fontsize: 100% # Base font size
# fontcolor: grey # Optional text color
highlight-style: pygments
format:
  html:
    embed-resources: true
    code-fold: true
    html-math-method: katex
  pdf:
    geometry:
      - top=30mm
      - left=30mm
  docx: default
editor:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
# Packages used in the project
required_packages <- c(
  "tidyverse",
  "shiny",
  "bslib"
  )

# Feature for automatically installing and downloading packages
load_or_install <- function(packages) {
  for (pkg in packages) {
    if (!require(pkg, character.only = TRUE)) {
      install.packages(pkg, dependencies = TRUE)
      library(pkg, character.only = TRUE)
    }
  }
}

load_or_install(required_packages)
```

## Development Guides

### General Guidelines

- All interface elements and outputs must be presented in Spanish.

- Outputs shown to users should be written in clear, easy-to-interpret text. In other words, prioritize human-readable explanations over technical descriptions.

- Prefer simple, readable, and reproducible code over unnecessarily complex solutions.

- At this stage, the priority is a functional application. Visual and aesthetic improvements will be implemented in future versions.

- All variable handling must occur within the local environment. Data curation and manipulation should be performed in this script immediately before each module or section.

- **Important:** Variable names must remain unchanged (immutable). However, they should be translated into human-readable Spanish concepts when presented outputs to users. 

- The use of the tidyverse ecosystem is preferred (functions, data manipulation, naming conventions, data structures, etc.).

### Additional Recommendations

#### Code Structure and Organization

- Remember the app is divided into logical modules. Pre-processing need to be dveloped in "Procesamiento de datos", your code inside the "Pestaña n".
- Use consistent naming conventions for objects and functions.
- Keep functions short and focused on a single task.
- Avoid code duplication; reuse functions whenever possible.

#### Reproducibility

- Ensure the script can run from start to finish without manual intervention.
- Set a seed when using random processes (`set.seed()`).
- Clearly state package dependencies (e.g., using `renv`, `DESCRIPTION`, or a requirements section in the README).

#### Collaboration and Version Control

- Document major decisions or assumptions within the code (comments).
- Keep code clean and remove unused variables or commented-out blocks before committing.

#### Documentation

- Comment code where necessary, especially in complex sections.
- Provide brief descriptions for your section in "Explanation Pestaña n"

#### Data Handling and Validation

- Validate inputs and handle missing or inconsistent data explicitly.
- Include basic error handling to prevent app crashes.
- Clearly define how missing values (`NA`) are treated.

#### User Experience (UX)

- Ensure outputs are intuitive and informative, even for non-technical users.
- Avoid overloading the interface with too much information at once.
- Use progressive disclosure: show detailed results only when needed.

#### Performance

- Avoid unnecessary recalculations inside reactive contexts.
- Use reactive expressions efficiently in Shiny.
- Optimize data manipulation steps.

### Optional (Highly Recommended)

- Follow a consistent style guide (e.g., tidyverse style guide).


## Explanation 

### Explanation Pestaña 1

### Explanation Pestaña 2

### Explanation Pestaña 3

### Explanation Pestaña 4

### Explanation Pestaña 5

## Loading datasets

```{r}
#| label: Load dataset

path = "Lupus_App_dataset.csv"
path_2 = "Personas_escaneadas_Lupus_Base_NeurolupusApp.csv"

load_lupus_data <- function(path) {
  raw_data <- read_csv(path, show_col_types = FALSE)
  
  curated_data <- raw_data %>%
    mutate(
      across(
        where(is.character),
        ~ na_if(., "")
      )
    )
  
  return(curated_data)
}

lupus_data <- load_lupus_data(path)
lupus_data_2 <- load_lupus_data(path_2)

```

## Formating datasets

```{r}
#| label: Formating dataset 01

formated_lupus_data <- lupus_data %>% 
  mutate(
    Disease.activity.classification = factor(
      Disease.activity.classification,
      levels = c(
        "No activity",
        "Mild activity",
        "Moderate activity",
        "High activity",
        "Very high activity"
      ),
      
      # Aplicamos la traducción directa al español
      labels = c(
        "Sin actividad", 
        "Leve", 
        "Moderada", 
        "Alta", 
        "Muy alta"
      ),
      
      ordered = TRUE
    ),
    
    
    clasi_ali = suppressWarnings(as.numeric(clasi_ali)),
    
    clasi_ali_es = factor(
      clasi_ali,
      levels = c(1, 2, 3),
      labels = c(
        "Hábitos inadecuados",
        "Hábitos parcialmente inadecuados",
        "Hábitos adecuados"
      ),
      ordered = TRUE
    ),
    
    comorbidities = factor(
      suppressWarnings(as.numeric(comorbidities)),
      levels = c(1,2,3,4,5,6,7,8,9,11),
      labels = c(
        "Diabetes mellitus",
        "Hipertensión",
        "Cáncer",
        "Enfermedad cardiovascular",
        "Osteoporosis",
        "Artritis reumatoide",
        "Esclerosis múltiple",
        "Problemas de tiroides",
        "Otra",
        "Ninguna"
      )
    ),
    
    home = factor(
      suppressWarnings(as.numeric(home)),
      levels = 1:32,
      labels = c(
        "Aguascalientes",
        "Baja California",
        "Baja California Sur",
        "Campeche",
        "Coahuila de Zaragoza",
        "Colima",
        "Chiapas",
        "Chihuahua",
        "Durango",
        "Ciudad de México",
        "Guanajuato",
        "Guerrero",
        "Hidalgo",
        "Jalisco",
        "México",
        "Michoacán de Ocampo",
        "Morelos",
        "Nayarit",
        "Nuevo León",
        "Oaxaca",
        "Puebla",
        "Querétaro",
        "Quintana Roo",
        "San Luis Potosí",
        "Sinaloa",
        "Sonora",
        "Tabasco",
        "Tamaulipas",
        "Tlaxcala",
        "Veracruz de Ignacio de la Llave",
        "Yucatán",
        "Zacatecas"
      )
    ),
    
    place_of_birth = factor(
      suppressWarnings(as.numeric(place_of_birth)),
      levels = 1:32,
      labels = c(
        "Aguascalientes",
        "Baja California",
        "Baja California Sur",
        "Campeche",
        "Coahuila de Zaragoza",
        "Colima",
        "Chiapas",
        "Chihuahua",
        "Durango",
        "Ciudad de México",
        "Guanajuato",
        "Guerrero",
        "Hidalgo",
        "Jalisco",
        "México",
        "Michoacán de Ocampo",
        "Morelos",
        "Nayarit",
        "Nuevo León",
        "Oaxaca",
        "Puebla",
        "Querétaro",
        "Quintana Roo",
        "San Luis Potosí",
        "Sinaloa",
        "Sonora",
        "Tabasco",
        "Tamaulipas",
        "Tlaxcala",
        "Veracruz de Ignacio de la Llave",
        "Yucatán",
        "Zacatecas"
      )
    ),
    
    school = factor(
      suppressWarnings(as.numeric(school)),
      levels = c(1, 2, 3, 4, 5, 6),
      labels = c(
        "Ninguno",
        "Primaria",
        "Secundaria",
        "Preparatoria/Carrera técnica",
        "Licenciatura",
        "Posgrado"
      )
    ),
    
    ocupation = factor(
      suppressWarnings(as.numeric(ocupation)),
      levels = c(1, 2, 3, 4),
      labels = c(
        "Estudiante",
        "Empleado",
        "Desempleado",
        "Jubilado/Retirado"
      )
    ),
    
    aditional_treatment = factor(
      suppressWarnings(as.numeric(aditional_treatment)),
      levels = c(1, 2),
      labels = c("Sí", "No")
    ),
    family_sle = factor(
  suppressWarnings(as.numeric(family_sle)),
  levels = c(0, 1),
  labels = c("Sí", "No")
    ),
  sex___1 = factor(
  suppressWarnings(as.numeric(sex___1)),
  levels = c(1, 0),
  labels = c("Mujer", "Hombre")
  ),
  prednisolona = factor(
  suppressWarnings(as.numeric(prednisolona)),
  levels = c(0,1,2,3,4,5,6,7,8),
  labels = c(
    "Prednisona",
    "Prednisolona",
    "Deflazacort",
    "Meticorten",
    "Metilprednisolona",
    "Calcort",
    "Betametazona",
    "No",
    "Otro"
  )
),

uses_corticosteroids = prednisolona != "No"
  )

treatment_labels <- c(
  "Antimaláricos",
  "Corticoesteroides",
  "Ácido micofenólico",
  "Azatioprina",
  "Metotrexato",
  "Rituximab",
  "Ciclosporina",
  "Ciclofosfamida",
  "Otros"
)

formated_lupus_data <- formated_lupus_data %>% 
  rowwise() %>% 
  mutate(
    treatment_summary = paste(
      treatment_labels[
        c_across(starts_with("treatment___")) == 1
      ],
      collapse = ", "
    ),
    
    treatment_summary = ifelse(
      treatment_summary == "",
      "Ninguno",
      treatment_summary
    ),
    
    num_treatments = sum(c_across(starts_with("treatment___"))),
    
    any_treatment = num_treatments > 0
  ) %>% 
  ungroup()

family_labels <- c(
  "Papá","Mamá","Herman@s","Abuel@",
  "Tí@","Prim@","Sobrin@","Hij@","Otro"
)

formated_lupus_data <- formated_lupus_data %>% 
  rowwise() %>% 
  mutate(
    # resumen de familiares con lupus
    family_member_sle_summary = if (family_sle == "Sí") {
      vals <- family_labels[
        c_across(starts_with("family_member_sle___")) == 1
      ]
      if (length(vals) == 0) "No especificado" else paste(vals, collapse = ", ")
    } else {
      NA_character_
    },
    
    # número de familiares
    num_family_members_sle = if (family_sle == "Sí") {
      sum(c_across(starts_with("family_member_sle___")))
    } else {
      NA_real_
    },
    
    # familiares de primer grado
    has_first_degree_sle = if (family_sle == "Sí") {
      sum(c_across(c(
        family_member_sle___0,
        family_member_sle___1,
        family_member_sle___2,
        family_member_sle___7
      ))) > 0
    } else {
      NA
    }
  ) %>% 
  ungroup()


formated_lupus_data <- formated_lupus_data %>% 
  mutate(
    # tratamiento total REAL (incluye corticoide)
    total_treatments = num_treatments + ifelse(uses_corticosteroids, 1, 0),
    
    treatment_profile = case_when(
      
      total_treatments == 0 ~ "Sin tratamiento",
      
      total_treatments == 1 & uses_corticosteroids ~ 
        paste0("Monoterapia: Corticoide (", prednisolona, ")"),
      
      total_treatments == 1 ~ 
        paste0("Monoterapia: ", treatment_summary),
      
      total_treatments > 1 ~ 
        paste0(
          "Politerapia: ",
          ifelse(
            treatment_summary == "Ninguno",
            paste0("Corticoide (", prednisolona, ")"),
            paste0(
              treatment_summary,
              ifelse(
                uses_corticosteroids,
                paste0(", Corticoide (", prednisolona, ")"),
                ""
              )
            )
          )
        )
    )
  )

glimpse(formated_lupus_data)
```

```{r}
#| label: Formatind Dataset 02
glimpse(lupus_data_2)
lupus_data_2 <- lupus_data_2 %>% 
  select(RedCapID,
         Grado_de_estudios,
         Anios_de_escolaridad,
         Puntaje_MoCA,
         MoCa_identificacion,
         MoCA_atencion,
         MoCA_lenguaje,
         MoCA_abstraccion,
         MoCA_recuerdo_diferido,
         MoCA_orientacion,
         Quality.control.T1,
         Quality.control.FLAIR,
         White.Matter..WM..volume.cm3,
         Total.lesion.count,
         Total.lesion.volume..absolute..cm3,
         Total.lesion.volume..normalized...,
         Total.lesion.burden,
         Caracteristicas_conservadas,
         Variantes_anatomicas,
         Aracnoide_selar,
         Quistes,
         Atrofia_cambios_involutivos,
         Gliosis_sustancia_blanca,
         Malacia_cortical,
         Lesiones_ocupantes_malformaciones
        
         )
glimpse(lupus_data_2)
formated_lupus_data_2 <- lupus_data_2 %>% 
  mutate(
    Grado_de_estudios = factor(
      Grado_de_estudios,
      levels = c(
        "Secundaria",
        "Preparatoria",
        "TSU",
        "Carrera_tecnica",
        "Licenciatura",
        "Maestria",
        "Posgrado",
        "Doctorado"),
      ordered = T
      ),
    
    Quality.control.T1 = factor(
      Quality.control.T1,
      levels = c(
        "A",
        "B",
        "C"),
      ordered = T
      ),
    Quality.control.FLAIR = factor(
      Quality.control.FLAIR,
      levels = c(
        "A",
        "B",
        "C"),
      ordered = T
      ),
    
    )

glimpse(formated_lupus_data_2)

```


## Reanaming dataset

```{r}
#| label: Renaming variables

formated_lupus_data <- formated_lupus_data %>%
   rename(Anios_viviendo_con_LES=dx_time,
          Anios_retraso_diagnostico=time_symptoms,
          Somatizacion=scl90r_score_som,
          Obsesion_y_compulsion=scl90r_score_oc,
          Sensitividad_interpersonal=scl90r_score_si,
          Depresion=scl90r_score_dep,
          Ansiedad=scl90r_score_ans,
          Hostilidad=scl90r_score_hos,
          Fobia=scl90r_score_fob,
          Ideacion_paranoide=scl90r_score_para,
          Psicoticismo=sci90r_score_psi,
          Tiene_familiar_con_LES=family_sle,
          Clasificacion_habitos_alimentacion=clasi_ali,
          Salud_mental=salud_mental_trans,
          Calidad_de_vida_dominio_psicologico=calc_psicologico,
          Calidad_de_vida_total=cal_totalcv,
          Comorbilidades=comorbidities,
          Tratamiento_adicional_a_LES=aditional_treatment,
          Tratamiento_corticoide=prednisolona,
          Nivel_socioeconomico=socioeconomic_status,
          Ocupacion=ocupation,
          Escolaridad=school,
          Estado_de_residencia=home,
          Edad=calculated_age,
          Indice_calidad_sueno=cal_glob,
          Dolor_corporal=dolor_corporal_trans,
          record_id=record_id,
          Sexo=sex___1,
          Estado_de_nacimiento=place_of_birth,
          Tratamiento_con_antimalaricos=treatment___0,
          Tratamiento_con_corticosteroides=treatment___1,
          Tratamiento_con_micofenolato=treatment___2,
          Tratamiento_con_azatioprina=treatment___3,
          Tratamiento_con_metotrexato=treatment___4,
          Tratamiento_con_rituximab=treatment___5,
          Tratamiento_con_ciclosporina=treatment___6,
          Tratamiento_con_ciclofosfamida=treatment___7,
          Tratamiento_con_otros_farmacos=treatment___8,
          Padre_con_LES=family_member_sle___0,
          Madre_con_LES=family_member_sle___1,
          Hermanos_con_LES=family_member_sle___2,
          Abuelos_con_LES=family_member_sle___3,
          Tios_con_LES=family_member_sle___4,
          Primos_con_LES=family_member_sle___5,
          Sobrinos_con_LES=family_member_sle___6,
          Hijos_con_LES=family_member_sle___7,
          Otros_familiares_con_LES=family_member_sle___8,
          Danio_ocular=Ocular,
          Danio_neuropsiquiatrico=Neuropsychiatric,
          Danio_renal=Renal,
          Danio_pulmonar=Pulmonary,
          Danio_cardiovascular=Cardiovascular,
          Danio_cardiomiopatia=Cardiomyopathy,
          Danio_vascular_periferico=Peripheral.vascular,
          Danio_gastrointestinal=Gastrointestinal,
          Danio_musculoesquletico=Musculoskeletal,
          Danio_cutaneo=Skin,
          Fallo_gonadal_prematuro=Premature.gonadal,
          Padece_diabetes=Diabetes,
          Presenta_cancer=Malignancy,
          Slicc=total.slicc,
          Sledai=total.sledai,
          Nivel_de_actividad_del_LES=Disease.activity.classification
)

formated_lupus_data
glimpse(formated_lupus_data)
```

## Joint Neurolupus dataset & variables selection

```{r}
#| label: Joint Neurolupus Dataset

# Unimos la base de datos de neuroimagen con los datos clínicos generales
formated_neurolupus_data <- formated_lupus_data_2 %>%
  left_join(
    formated_lupus_data, 
    by = c("RedCapID" = "record_id") # Igualamos las llaves identificadoras
  ) %>%
  # Opcional: Reordenar para tener el ID al inicio y limpiar columnas redundantes si las hubiera
  relocate(RedCapID)

# Verificamos la integración correcta de las variables
glimpse(formated_neurolupus_data)

# Seleccionamos variables

formated_neurolupus_data_02 <- formated_neurolupus_data %>% 
  select(Puntaje_MoCA,
         MoCA_atencion,
         MoCa_identificacion,
         MoCA_lenguaje,
         MoCA_abstraccion,
         MoCA_recuerdo_diferido,
         MoCA_orientacion,
         Total.lesion.count,
         Total.lesion.burden,
         Caracteristicas_conservadas,
         Variantes_anatomicas,
         Aracnoide_selar,
         Quistes,
         Atrofia_cambios_involutivos,
         Gliosis_sustancia_blanca,
         Malacia_cortical,
         Lesiones_ocupantes_malformaciones,
         Anios_viviendo_con_LES,
         Anios_retraso_diagnostico,
         Anios_de_escolaridad,
         Somatizacion,
         Obsesion_y_compulsion,
         Sensitividad_interpersonal,
         Depresion,
         Ansiedad,
         Hostilidad,
         Fobia,
         Ideacion_paranoide,
         Psicoticismo,
         Salud_mental,
         Calidad_de_vida_dominio_psicologico,
         Calidad_de_vida_total,
         Nivel_socioeconomico,
         Edad,
         Indice_calidad_sueno,
         Dolor_corporal,
         Sexo,
         Tratamiento_con_antimalaricos,
         uses_corticosteroids,
         Tratamiento_con_ciclofosfamida,
         Tratamiento_con_rituximab,
         has_first_degree_sle,
         num_family_members_sle,
         num_treatments,
         any_treatment,
         Danio_neuropsiquiatrico,
         Danio_cardiovascular,
         Danio_vascular_periferico,
         Slicc,
         Sledai
         )
glimpse(formated_neurolupus_data_02)

```


## Defining Human-readable dictionaries

```{r}
#| label: Neurolupus dictionary

# Diccionario completo para el módulo de Neurolupus
diccionario_neuro_completo <- list(
  
  "Evaluación Cognitiva (MoCA)" = c(
    "Puntaje Total MoCA" = "Puntaje_MoCA",
    "Atención" = "MoCA_atencion",
    "Identificación" = "MoCa_identificacion",
    "Lenguaje" = "MoCA_lenguaje",
    "Abstracción" = "MoCA_abstraccion",
    "Recuerdo Diferido" = "MoCA_recuerdo_diferido",
    "Orientación" = "MoCA_orientacion"
  ),
  
  "Hallazgos de Neuroimagen (RM)" = c(
    "Conteo Total de Lesiones" = "Total.lesion.count",
    "Carga Lesional (Burden)" = "Total.lesion.burden",
    "Características Conservadas" = "Caracteristicas_conservadas",
    "Variantes Anatómicas" = "Variantes_anatomicas",
    "Aracnoide Selar" = "Aracnoide_selar",
    "Quistes" = "Quistes",
    "Atrofia / Cambios Involutivos" = "Atrofia_cambios_involutivos",
    "Gliosis en Sustancia Blanca" = "Gliosis_sustancia_blanca",
    "Malacia Cortical" = "Malacia_cortical",
    "Lesiones Ocupantes / Malformaciones" = "Lesiones_ocupantes_malformaciones"
  ),
  
  "Salud Mental y Psicológica (SCL-90-R y PROMs)" = c(
    "Percepción Global de Salud Mental" = "Salud_mental",
    "Somatización" = "Somatizacion",
    "Obsesión y Compulsión" = "Obsesion_y_compulsion",
    "Sensitividad Interpersonal" = "Sensitividad_interpersonal",
    "Depresión" = "Depresion",
    "Ansiedad" = "Ansiedad",
    "Hostilidad" = "Hostilidad",
    "Fobia" = "Fobia",
    "Ideación Paranoide" = "Ideacion_paranoide",
    "Psicoticismo" = "Psicoticismo",
    "Calidad de Vida (Dominio Psicológico)" = "Calidad_de_vida_dominio_psicologico",
    "Calidad de Vida Total" = "Calidad_de_vida_total",
    "Índice de Calidad de Sueño" = "Indice_calidad_sueno",
    "Dolor Corporal" = "Dolor_corporal"
  ),
  
  "Actividad y Daño Acumulado" = c(
    "Puntaje SLEDAI (Actividad)" = "Sledai",
    "Puntaje SLICC (Daño Total)" = "Slicc",
    "Daño Neuropsiquiátrico" = "Danio_neuropsiquiatrico",
    "Daño Cardiovascular" = "Danio_cardiovascular",
    "Daño Vascular Periférico" = "Danio_vascular_periferico"
  ),
  
  "Perfil Clínico y Tratamiento" = c(
    "Años viviendo con LES" = "Anios_viviendo_con_LES",
    "Años de Retraso Diagnóstico" = "Anios_retraso_diagnostico",
    "Número Total de Tratamientos" = "num_treatments",
    "Tiene Tratamiento Activo" = "any_treatment",
    "Uso de Corticoesteroides" = "uses_corticosteroids",
    "Tratamiento con Antimaláricos" = "Tratamiento_con_antimalaricos",
    "Tratamiento con Ciclofosfamida" = "Tratamiento_con_ciclofosfamida",
    "Tratamiento con Rituximab" = "Tratamiento_con_rituximab"
  ),
  
  "Sociodemográficos y Antecedentes" = c(
    "Edad" = "Edad",
    "Sexo" = "Sexo",
    "Nivel Socioeconómico" = "Nivel_socioeconomico",
    "Años de Escolaridad" = "Anios_de_escolaridad",
    "Familiar de 1er Grado con LES" = "has_first_degree_sle",
    "Número de Familiares con LES" = "num_family_members_sle"
  )
)

```


## Shiny App

```{r}
#| label: Lupus App

# Procesamiento de datos

# Interfaz de Usuario (UI)
ui <- page_navbar(
  title = "LupusRGMX Data App",
  theme = bs_theme(version = 5, bootswatch = "flatly"),
  
  # Pestaña 1: Reporte General
  nav_panel(title = "Reporte General",
            sidebarLayout(
              sidebarPanel(
                h4("Filtros Generales"),
                # Aquí irán los selectInput para Edad, Sexo, SLEDAI, etc.
                selectInput("var_gen", "Selecciona variables:", choices = c("Edad", "Sexo", "Comorbilidades"), multiple = TRUE),
                actionButton("run_gen", "Generar Reporte", class = "btn-primary")
              ),
              mainPanel(
                h3("Resumen Descriptivo"),
                textOutput("txt_reporte_gen") # Salida de texto reactiva
              )
            )
  ),
  
  # Pestaña 2: Modelado Estadístico
  nav_panel(title = "Modelado Estadístico",
            sidebarLayout(
              sidebarPanel(
                h4("Configuración del Modelo"),
                selectInput("var_target", "Variable Objetivo (Target):", choices = c("SLEDAI", "Calidad de Vida", "Nefritis Lúpica")),
                selectInput("var_predict", "Variables Predictoras (Features):", choices = c("Uso de Corticoesteroides", "Tiempo de diagnóstico"), multiple = TRUE),
                radioButtons("modelo_tipo", "Tipo de Modelo:", choices = c("Regresión Lineal", "Regresión Binomial", "Chi-cuadrada")),
                actionButton("run_model", "Ejecutar Modelo", class = "btn-primary")
              ),
              mainPanel(
                h3("Resultados del Modelo"),
                textOutput("txt_reporte_modelo") # Salida de texto reactiva
              )
            )
  ),
  
  # Pestaña 3: Neurolupus
  nav_panel(title = "Neurolupus",
            # En la sección de la UI -> nav_panel(title = "Neurolupus", ...)
sidebarLayout(
  sidebarPanel(
    h4("Análisis de Predicción Neurolupus"),
    # Selección de Variable Objetivo
    selectInput("neuro_target", "Variable Objetivo (Resultado):", 
                choices = diccionario_neuro_completo),
    
    # Selección de Predictores (Máximo 5)
    selectizeInput("neuro_predictors", "Variables Predictoras (Máximo 5):", 
                   choices = diccionario_neuro_completo, 
                   multiple = TRUE, 
                   options = list(maxItems = 5)),
    
    hr(),
    actionButton("run_neuro_model", "Efectuar Análisis Estadístico", class = "btn-primary")
  ),
  mainPanel(
    h3("Reporte de Análisis Neurocognitivo"),
    wellPanel(
      verbatimTextOutput("txt_reporte_neuro_model")
    )
  )
)
  ),
  
  
  # Pestaña 4: Expresión Génica
  nav_panel(title = "Expresión Génica",
            sidebarLayout(
              sidebarPanel(
                h4("Análisis Transcriptómico"),
                selectInput("var_gene", "Selecciona Gene(s):", 
                            choices = c("IFNIT", "STAT1", "OAS1", "MX1"), 
                            multiple = TRUE),
                radioButtons("plot_type", "Tipo de visualización:", 
                             choices = c("Boxplot de Expresión", "Volcano Plot", "Heatmap")),
                actionButton("run_gene", "Analizar Expresión", class = "btn-primary")
              ),
              mainPanel(
                h3("Resultados de Expresión Diferencial"),
                textOutput("txt_reporte_gene"), # Salida de texto
                plotOutput("plot_gene")        # Espacio para gráfica futura
              )
            )
  ),
  
  # Pestaña 5: Acceso a Datos
  nav_panel(title = "Acceso a Datos Crudos",
            div(class = "container mt-5",
                h2("Solicitud de Datos"),
                p("Los datos crudos de este estudio se encuentran resguardados para proteger la privacidad de las y los participantes."),
                p("Si eres investigador y deseas acceder a la base de datos para análisis colaborativos, por favor completa el formulario de solicitud formal."),
                br(),
                a(href = "URL_DE_TU_REDCAP", target = "_blank", class = "btn btn-lg btn-success", "Ir al Formulario en RedCap")
            )
  )
)

# Lógica del Servidor (Server)
server <- function(input, output, session) {
  
  # Lógica Pestaña 1
  observeEvent(input$run_gen, {
    output$txt_reporte_gen <- renderText({
      paste("Reporte generado para las variables:", paste(input$var_gen, collapse = ", "), 
            ". Aquí se insertará la lógica para calcular y redactar promedios, conteos, etc.")
    })
  })
  
  # Lógica Pestaña 2
  observeEvent(input$run_model, {
    output$txt_reporte_modelo <- renderText({
      paste("Ejecutando", input$modelo_tipo, "para predecir", input$var_target, 
            "utilizando", paste(input$var_predict, collapse = ", "), 
            ". Aquí se interpretarán los coeficientes y el p-value en formato texto.")
    })
  })
  
  # Lógica Pestaña 3
  # Dentro de la función server
    # Dentro de la función server
  observeEvent(input$run_neuro_model, {
    output$txt_reporte_neuro_model <- renderText({
      req(input$neuro_target, input$neuro_predictors)
      
      # 1. Preparación de datos y etiquetas
      df_model <- formated_neurolupus_data_02 
      target_var <- input$neuro_target
      pred_vars <- input$neuro_predictors
      
      # Recuperar nombres legibles para el reporte
      all_labels <- unlist(diccionario_neuro_completo)
      target_label <- names(all_labels)[all_labels == target_var]
      
      # 2. Detección automática del tipo de modelo
      is_binary <- is.factor(df_model[[target_var]]) || length(unique(na.omit(df_model[[target_var]]))) == 2
      model_type <- if(is_binary) "Regresión Binomial (Logística)" else "Regresión Lineal Múltiple"
      
      # Construcción de fórmula
      formula_model <- as.formula(paste(target_var, "~", paste(pred_vars, collapse = " + ")))
      
      # 3. Ejecución del Modelo
      tryCatch({
        if(is_binary) {
          fit <- glm(formula_model, data = df_model, family = "binomial")
          stats <- summary(fit)$coefficients
        } else {
          fit <- lm(formula_model, data = df_model)
          stats <- summary(fit)$coefficients
          r_sq <- round(summary(fit)$r.squared * 100, 1)
        }
        
        # 4. Redacción del Reporte Humano (almacenando línea por línea)
        reporte_lineas <- c(
          "### RESUMEN DEL MODELO ESTRUCTURAL ###",
          "",
          paste("Objetivo: Evaluar el efecto sobre '", target_label, "'.", sep = ""),
          paste("Método: Se aplicó una", model_type, "ajustada por", length(pred_vars), "variables."),
          ""
        )
        
        # Evaluación de predictores significativos (p < 0.05)
        sig_rows <- rownames(stats)[stats[,4] < 0.05 & rownames(stats) != "(Intercept)"]
        
        if(length(sig_rows) > 0) {
          reporte_lineas <- c(reporte_lineas, "RESULTADOS CLAVE:")
          
          for(row in sig_rows) {
            clean_label <- names(all_labels)[all_labels == row]
            p_val <- round(stats[row, 4], 4)
            
            if(is_binary) {
              # Calcular Odds Ratio para modelos binomiales
              or_val <- round(exp(stats[row, 1]), 2)
              linea_sig <- paste("- La variable '", clean_label, "' es significativa (p = ", p_val, "). Por cada unidad de incremento, la probabilidad del evento aumenta ", or_val, " veces (Odds Ratio).", sep = "")
              reporte_lineas <- c(reporte_lineas, linea_sig)
              
            } else {
              coeff <- round(stats[row, 1], 2)
              linea_sig <- paste("- '", clean_label, "' muestra una asociación significativa (p = ", p_val, "). Un aumento en esta variable se traduce en un cambio de ", coeff, " unidades en '", target_label, "'.", sep = "")
              reporte_lineas <- c(reporte_lineas, linea_sig)
            }
          }
        } else {
          reporte_lineas <- c(reporte_lineas, "RESULTADOS: No se detectaron asociaciones con significancia estadística (p < 0.05) para los predictores seleccionados en esta muestra.")
        }
        
        if(!is_binary) {
          reporte_lineas <- c(reporte_lineas, "", paste("AJUSTE: El modelo explica el", r_sq, "% de la variabilidad observada."))
        }
        
        # Unimos todas las líneas con saltos de línea (\n)
        return(paste(reporte_lineas, collapse = "\n"))
        
      }, error = function(e) {
        msg_error <- c(
          "Error en el procesamiento: Los datos seleccionados no permiten la convergencia del modelo.",
          "Verifique que las variables tengan suficientes casos válidos (N).",
          paste("Detalle:", e$message)
        )
        return(paste(msg_error, collapse = "\n"))
      })
    })
  }) 
  
   # Lógica Pestaña 4
    observeEvent(input$run_gene, {
      output$txt_reporte_gene <- renderText({
        paste("Analizando la expresión de los genes:", paste(input$var_gene, collapse = ", "), 
              "mediante un", input$plot_type, 
              ". En esta sección se integrará el análisis de cuentas normalizadas o Fold-Change.")
      })
      
      # Placeholder para la gráfica
      output$plot_gene <- renderPlot({
        # Aquí irá la lógica de ggplot2 una vez se carguen los datos de expresión
      })
    })
    
    
  }


shinyApp(ui, server)

```




